European or American? A re-examination of the transatlantic influence over the IMF

### 2025-07-29 manuscript/manuscript.qmd
EU lit review revised_0729

---
title: European or American? A re-examination of the transatlantic influence over the IMF
date: last-modified
authors:
  - name: Dianyi Yang
    email: d.yang17@lse.ac.uk
    orcid: 0009-0004-4652-3429
    correspondence: true
  - name: Tong Hu
    orcid: 0009-0003-5234-8164
format:
  #native: default
  #docx: default
  orcid-pdf:
    keep-tex: true
    number-sections: true
    include-in-header:
      - text: |
          \newcommand{\centerstart}{\begin{center}}
          \newcommand{\centerend}{\end{center}}
          \usepackage{amsthm, thm-restate}
          \newtheorem{hyp}{Hypothesis}
          \newtheorem{subhyp}{Hypothesis}[hyp]
          \renewcommand{\thesubhyp}{\thehyp\alph{subhyp}}
          \usepackage{chngcntr}
          \usepackage{apptools}
          \AtAppendix{\counterwithin{figure}{section}}
abstract: This is an abstract.
keywords: IMF, US hegemony, Europe
execute:
  echo: false
  warning: false
  error: false
  message: false
---

@@

Third, European influence over IMF lending has often served to advance strategic financial and geopolitical interests. This was particularly evident during the Eurozone crisis, as mentioned earlier, when major European economies—most notably Germany and France—sought to mitigate systemic risks arising from their banking sector’s exposure to highly indebted member states. Fiscal surveillance mechanisms, such as the EU’s 3% deficit ceiling, were complemented by IMF-backed assistance programs aimed at restoring financial stability [@heimberger2025, p. 9]. The initiation of the IMF’s financial assistance to Greece, for example, was undertaken at Germany’s explicit request [@kickert2019]. Beyond economic stabilization, European policymakers also leveraged IMF involvement to enhance the legitimacy of domestic austerity programs, using the Fund’s reputation as an external technocratic authority to depoliticize painful fiscal adjustments [@kickert2019].

>>>>

---
title: European or American? A re-examination of the transatlantic influence over the IMF
date: last-modified
authors:
  - name: Dianyi Yang
    email: d.yang17@lse.ac.uk
    orcid: 0009-0004-4652-3429
    correspondence: true
  - name: Tong Hu
    orcid: 0009-0003-5234-8164
  - name: Zhenyi Chen
    orcid: 0009-0008-2910-5187
format:
  #native: default
  #docx: default
  orcid-pdf:
    keep-tex: true
    number-sections: true
    include-in-header:
      - text: |
          \newcommand{\centerstart}{\begin{center}}
          \newcommand{\centerend}{\end{center}}
          \usepackage{amsthm, thm-restate}
          \newtheorem{hyp}{Hypothesis}
          \newtheorem{subhyp}{Hypothesis}[hyp]
          \renewcommand{\thesubhyp}{\thehyp\alph{subhyp}}
          \usepackage{chngcntr}
          \usepackage{apptools}
          \AtAppendix{\counterwithin{figure}{section}}
abstract: This is an abstract.
keywords: IMF, US hegemony, Europe
execute:
  echo: false
  warning: false
  error: false
  message: false
---

@@

Third, European influence over IMF lending has often served to advance strategic financial and geopolitical interests. This was particularly evident during the Eurozone crisis, as mentioned earlier, when major European economies—most notably Germany and France—sought to mitigate systemic risks arising from their banking sector’s exposure to highly indebted member states. Fiscal surveillance mechanisms, such as the EU’s 3% deficit ceiling, were complemented by IMF-backed assistance programs aimed at restoring financial stability [@heimberger2025, p. 9]. The initiation of the IMF’s financial assistance to Greece, for example, was undertaken at Germany’s explicit request [@kickert2019a]. Beyond economic stabilization, European policymakers also leveraged IMF involvement to enhance the legitimacy of domestic austerity programs, using the Fund’s reputation as an external technocratic authority to depoliticize painful fiscal adjustments [@kickert2019a].
